---
title: "In vitro bacterial dynamics"
subtitle: "Effect of antifungal on Pe and two common bacteria from Drosophila's microbiota"
date: ""
chunk_output_type: inline
output: html_document
theme: yeti
toc: yes
toc_float: yes
authors: "Youn"
---
  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
```


# Packages

```{r packages-loading}
#[R v4.0.5]

if(require("pacman")==T){
  library("pacman")
}else{
  install.packages("pacman")
}

pacman::p_load(Rmisc, tidyverse, tidybayes, ggplot2, kableExtra, survival, survminer, performance, car, lme4, multcomp, emmeans, brms, cmdstanr, ggtext, relaimpo, gridExtra, bayesplot, flextable)


library(here)      # CRAN v1.0.1
## tool for workspace location management
## automatically detects the workspace after loading the package if Rstudio was opened using the script file (or with new R project)
## load last to avoid conflicts
here()

```

# Preprocess

Loading the data

```{r data-load}
plating_data=read.table(here("data_CFU.txt"),h=T, dec=".")
head(plating_data)

```

Editing the data table for cleaner graphical outputs

```{r data tinkering}
data1 = plating_data %>%
  mutate(time = as.factor(time)) %>%
  mutate(species = str_replace_all(species, "P_entomophila", "*P. entomophila*")) %>%
  mutate(species = str_replace_all(species, "A_pasterianus", "*A. pasterianus*")) %>%
  mutate(species = str_replace_all(species, "L_plantarum", "*L. plantarum*")) %>%
  mutate(antifungal = str_replace_all(antifungal, "nipagin", "methylparaben")) %>%
    mutate(antifungal = str_replace_all(antifungal, "propionic", "propionic acid")) %>%
  mutate(species = factor(species, levels=c("*A. pasterianus*", "*L. plantarum*", "*P. entomophila*"))) %>%
  mutate(treatment = factor(treatment, levels=c("control", "antifungal"))) %>%
  arrange(species) %>%
  arrange(treatment)


data2 = data1 %>%
  mutate(CFU = ifelse(treatment == "antifungal" & time == 24, NA, CFU))
data2=subset(data1, data1$species=="*P. entomophila*")
data2=subset(data2, data2$antifungal=="methylparaben")
data2 = data2 %>%
  mutate(treatment = str_replace_all(treatment, "control", "Without mp")) %>%
  mutate(treatment = str_replace_all(treatment, "antifungal", "With mp"))

data3=subset(data1, data1$species=="*P. entomophila*")
data3=subset(data3, data3$antifungal=="propionic acid")
data3 = data3 %>%
  mutate(treatment = str_replace_all(treatment, "control", "Without <br> propionic acid")) %>%
  mutate(treatment = str_replace_all(treatment, "antifungal", "With <br> propionic acid"))


mycol=c("#FFC107","#8A68D7", "#f01010")
mycol2=c("#f01010", "#ffffff")
mycol3=c("#f01010", "#f01010")


```

# Plots of bacterial loads

Complete plot

```{r plating plot}
##################################################################
### Plotting abundance dynamics for each vial (N=6 mean of two plated flies for each time)

plot_cfu2=ggplot(aes(x=time, y=(CFU+1), fill=species, colour=species), data=data1)+
  stat_summary(fun.data = "mean_cl_boot", fun.args = list(conf.int = .95), geom="point", aes(group = interaction(species,treatment),  shape = treatment), position=position_dodge(0), alpha=0.5, size = 2, show.legend = T)+
  stat_summary(fun.data = "mean_cl_boot", fun.args = list(conf.int = .95), geom="linerange", aes(group = interaction(species,treatment)), position=position_dodge(0), alpha=1, size = 0.3, show.legend = F)+
  stat_summary(fun.data = "mean_cl_boot", geom="line", aes(group = interaction(species,treatment), linetype = treatment), position=position_dodge(0), alpha=1, size = 0.3, show.legend = F)+
  scale_fill_manual(values=mycol)+
  scale_color_manual(values=mycol)+
  scale_shape_manual(values=c(16,17))+
  scale_alpha_manual(values=c(1,0.7,0.4))+
  labs(x= "Time (hours)", y="CFU per mL")+
  annotate("text", x = 0.65, y = 6666, label = "min.", vjust = -0.5, size= 2)+
  geom_hline(yintercept = 6666, linetype='dotted', color = 'black')+
  coord_cartesian(ylim = c(1,4000000000), clip="off")+
  facet_grid(antifungal~species)+
  scale_y_log10(breaks = c(1, 100, 10000, 1000000, 100000000), labels=scales::scientific)+
  theme_bw(base_size = 12)+
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(color = "grey95", size = 0.4), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(angle=0, vjust=0, hjust=0.5), axis.title.y = element_markdown(), legend.position = "right", legend.text = element_markdown(), strip.text.x = element_markdown(), strip.text.y = element_markdown(), legend.title = element_markdown())

plot_cfu2

pdf(height=4, width=6, file=here("plot_CFU_antifungal.pdf"))
plot_cfu2+
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(color = "grey95", size = 0.4), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(size = 8, angle=0, vjust=0, hjust=0.5), axis.text.y = element_text(size = 8), axis.title.x = element_markdown(size = 10, padding = margin(5, 0, 0, 0)), axis.title.y = element_markdown(size = 10), legend.position = "right", legend.text = element_markdown(size = 8), strip.text.x = element_blank(), strip.background.x = element_blank(), strip.text.y = element_markdown(), strip.background.y = element_blank(), legend.title = element_blank(), plot.margin = margin(t = 30, r = 22, b = 6, l = 6, unit = "pt"), panel.border = element_rect(colour = "grey20", fill=NA, size=0.4))
dev.off()


```

# Plot with only Pe
methylparaben

```{r plating plot only Pe methylparaben}
##################################################################
### Plotting abundance dynamics for each vial (N=6 mean of two plated flies for each time)

annotation_df1 <- data.frame(
  treatment = c('With mp', 'Without mp'),
  x = c(0.3, 0.3),
  y = c(15000, 15000),
  label = 'min.'
)
annotation_df2 <- data.frame(
  treatment = c('With mp', 'Without mp'),
  x = c(0.3, 0.3),
  y = c(8000000000, 8000000000)
)
annotation_df3 <- data.frame(
  treatment = 'With mp',
  x = 2,
  y = 1000,
  label = 'below <br> detection'
)

plot_cfu3=ggplot(aes(x=time, y=(CFU+1), fill=treatment, color=treatment), data=data2)+
  stat_summary(fun.data = "mean_cl_boot", geom="line", aes(group = interaction(species,treatment), linetype = treatment), position=position_dodge(0), alpha=1, size = 0.3, show.legend = F)+
  geom_dotplot(binaxis='y',binwidth = 0.30, stackdir='center', position=position_dodge(0), dotsize = 1, show.legend = F)+
  scale_fill_manual(values=mycol2)+
  scale_color_manual(values=mycol3)+
  scale_linetype_manual(values=c(2,1))+
  labs(title="", x= "Time (hours)", y="CFU per mL")+
  geom_richtext(data = annotation_df1, aes(x = x, y = y, label = label), color = 'black', hjust = 0, vjust = 0.5, size = 2.5, show.legend = FALSE, fill = NA, label.color = NA) +
  geom_richtext(data = annotation_df2, aes(x = x, y = y, label = treatment), color = 'black', hjust = 0, vjust = 1, size = 3, show.legend = FALSE, fill = NA, label.color = NA) +
  geom_richtext(data = annotation_df3, aes(x = x, y = y, label = label), color = 'black', hjust = 0.5, vjust = 0.5, size = 2.5, show.legend = FALSE, fill = NA, label.color = NA) +
  geom_hline(yintercept = 6666, linetype='dotted', color = 'black')+
  coord_cartesian(ylim = c(1,4000000000), clip="off")+
  scale_y_log10(breaks = c(1, 100, 10000, 1000000, 100000000), labels=scales::scientific)+
  facet_grid(~treatment)+
  theme_bw(base_size = 12)+
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(color = "grey95", size = 0.4), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(angle=0, vjust=0, hjust=0.5), axis.title.y = element_markdown(), legend.position = "right", legend.text = element_markdown(), strip.text.x = element_markdown(), strip.text.y = element_markdown(), legend.title = element_markdown())

plot_cfu3

pdf(height=3, width=3.5, file=here("plot_CFU_methylparaben_Pe.pdf"))
plot_cfu3+
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(color = "grey95", size = 0.4), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(size = 8, angle=0, vjust=0, hjust=0.5), axis.text.y = element_text(size = 8), axis.title.x = element_markdown(size = 12, padding = margin(5, 0, 0, 0)), axis.title.y = element_markdown(size = 12), legend.position = "right", legend.text = element_markdown(size = 8), strip.text.x = element_blank(), strip.background.x = element_blank(), strip.text.y = element_markdown(), strip.background.y = element_blank(), plot.title = element_markdown(size=8), legend.title = element_blank(), plot.margin = margin(t = 10, r = 22, b = 6, l = 6, unit = "pt"), panel.border = element_rect(colour = "grey20", fill=NA, size=0.4))
dev.off()


```
propionic acid

```{r plating plot only Pe propionic}
##################################################################
### Plotting abundance dynamics for each vial (N=6 mean of two plated flies for each time)

annotation_df1 <- data.frame(
  treatment = c('With <br> propionic acid', 'Without <br> propionic acid'),
  x = c(0.3, 0.3),
  y = c(15000, 15000),
  label = 'min.'
)
annotation_df2 <- data.frame(
  treatment = c('With <br> propionic acid', 'Without <br> propionic acid'),
  x = c(0.3, 0.3),
  y = c(8000000000, 8000000000)
)
annotation_df3 <- data.frame(
  treatment = 'With <br> propionic acid',
  x = 2,
  y = 1000,
  label = 'below <br> detection'
)


plot_cfu4=ggplot(aes(x=time, y=(CFU+1), fill=treatment, color=treatment), data=data3)+
  stat_summary(fun.data = "mean_cl_boot", geom="line", aes(group = interaction(species,treatment), linetype = treatment), position=position_dodge(0), alpha=1, size = 0.3, show.legend = F)+
  geom_dotplot(binaxis='y',binwidth = 0.30, stackdir='center', position=position_dodge(0), dotsize = 1, show.legend = F)+
  scale_fill_manual(values=mycol2)+
  scale_color_manual(values=mycol3)+
  scale_linetype_manual(values=c(2,1))+
  labs(title="", x= "Time (hours)", y="CFU per mL")+
  geom_richtext(data = annotation_df1, aes(x = x, y = y, label = label), color = 'black', hjust = 0, vjust = 0.5, size = 2.5, show.legend = FALSE, fill = NA, label.color = NA) +
  geom_richtext(data = annotation_df2, aes(x = x, y = y, label = treatment), color = 'black', hjust = 0, vjust = 1, size = 3, show.legend = FALSE, fill = NA, label.color = NA) +
  geom_richtext(data = annotation_df3, aes(x = x, y = y, label = label), color = 'black', hjust = 0.5, vjust = 0.5, size = 2.5, show.legend = FALSE, fill = NA, label.color = NA) +
  geom_hline(yintercept = 6666, linetype='dotted', color = 'black')+
  coord_cartesian(ylim = c(1,4000000000), clip="off")+
  scale_y_log10(breaks = c(1, 100, 10000, 1000000, 100000000), labels=scales::scientific)+
  facet_grid(~treatment)+
  theme_bw(base_size = 12)+
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(color = "grey95", size = 0.4), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(angle=0, vjust=0, hjust=0.5), axis.title.y = element_markdown(), legend.position = "right", legend.text = element_markdown(), strip.text.x = element_markdown(), strip.text.y = element_markdown(), legend.title = element_markdown())

plot_cfu4

pdf(height=3, width=3.5, file=here("plot_CFU_propionic_Pe.pdf"))
plot_cfu4+
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(color = "grey95", size = 0.4), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(size = 8, angle=0, vjust=0, hjust=0.5), axis.text.y = element_text(size = 8), axis.title.x = element_markdown(size = 12, padding = margin(5, 0, 0, 0)), axis.title.y = element_markdown(size = 12), legend.position = "right", legend.text = element_markdown(size = 8), strip.text.x = element_blank(), strip.background.x = element_blank(), strip.text.y = element_markdown(), strip.background.y = element_blank(), plot.title = element_markdown(size=8), legend.title = element_blank(), plot.margin = margin(t = 10, r = 22, b = 6, l = 6, unit = "pt"), panel.border = element_rect(colour = "grey20", fill=NA, size=0.4))
dev.off()

```

# Analysis

Bayesian model of the effect of time and food preservative on bacterial growth of 3 species

```{r model construction}

if (file.exists(here("mod_brms_cfu.Rdata"))) {
  load(here("mod_brms_cfu.Rdata"))
} else {
mod_brms_cfu <- brm(
  bf(log(CFU+1) ~ treatment+time+treatment:time+(1|replicate)), 
    data = data2,
    chains = 4, iter = 10000, warmup = 2000, seed = 1312,
    prior = c(
      set_prior("normal(0,5)", class = "b"),
      set_prior("normal(0,1)", class = "sd")
    ),
  backend = "cmdstanr"
)
save(
  list = c("mod_brms_cfu"),
  file = here("mod_brms_cfu.Rdata")
)
}
```

Check of model sanity and making tables with fitted

```{r extracting posteriors}
pp_check(mod_brms_cfu)
pp_check(mod_brms_cfu, "stat_2d")

fits_bay1 <- data2 |> 
  dplyr::select(species, antifungal, treatment, replicate, time, CFU) |> 
  distinct()  |> 
  add_epred_draws(mod_brms_cfu) |> 
  ungroup()

fits_bay1
```

Plot of observed vs fitted values

```{r obs vs fitted plot}
Plot_CFU_obsvsfitted=ggplot(aes(x=time, y=(CFU+1), fill=species, colour=species), data=data2)+
  stat_eye(aes(x=time, y=(exp(.epred)), group = interaction(species,treatment)), normalize = "xy",.width = c(0.001, 0.95), shape = "", point_size = 1, point_color = "black", slab_fill = "black", slab_alpha = 0.2, interval_alpha = 0, data=fits_bay1, show.legend = F)+
  stat_summary(fun.data = "mean_cl_boot", fun.args = list(conf.int = .95), geom="point", aes(group = interaction(species,treatment),  shape = treatment), position=position_dodge(0), alpha=0.5, size = 3, show.legend = F)+
  stat_summary(fun.data = "mean_cl_boot", fun.args = list(conf.int = .95), geom="linerange", aes(group = interaction(species,treatment)), position=position_dodge(0), alpha=1, size = 0.3, show.legend = F)+
  stat_summary(fun.data = "mean_cl_boot", geom="line", aes(group = interaction(species,treatment), linetype = treatment), position=position_dodge(0), alpha=1, size = 0.3, show.legend = F)+
  scale_fill_manual(values=mycol2)+
  scale_color_manual(values=mycol2)+
  scale_shape_manual(values=c(16,17))+
  labs(x= "Time (hours)", y="CFU per mL")+
  facet_grid(~species)+
  annotate("text", x = 0.65, y = 6666, label = "min.", vjust = -0.5, size = 2)+
  geom_hline(yintercept = 6666, linetype='dotted', color = 'black', size = 0.3)+
  coord_cartesian(ylim = c(1,4000000000), clip="off")+
  scale_y_log10(breaks = c(1, 100, 10000, 1000000, 100000000), labels=scales::scientific)+
  theme_bw(base_size = 12)+
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(color = "grey95", size = 0.4), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(angle=0, vjust=0, hjust=0.5), axis.title.y = element_markdown(), legend.position = "right", legend.text = element_markdown(), strip.text.x = element_markdown(), strip.text.y = element_markdown(), legend.title = element_markdown())

Plot_CFU_obsvsfitted
```

Extracting model fitted into two tables

```{r extract difference table from fits}

fits_bay1=subset(fits_bay1, fits_bay1$replicate==1)
fits_bay1$exp.epred=exp(fits_bay1$.epred)

table_pred <-flextable(fits_bay1 |>
  mutate(.iteration = .draw) |>
  group_by(time) |>
  compare_levels(exp.epred, by = treatment, comparison = control) |>
  mean_hdi(exp.epred) |> ## here with mean, but it also works with mean_hdi, median_qi, mean_qi
  mutate(.lower = round(.lower, digits = 2)) |>
  mutate(.upper = round(.upper, digits = 2)) |>
  mutate(exp.epred = round(exp.epred, digits = 2)) |>
  dplyr::select(treatment, time, exp.epred, .lower, .upper) |>
  rename("LowerCI" = ".lower", "Prediction" = "exp.epred", "UpperCI" = ".upper", "Treatment" = "treatment", "Time" = "time")) |>
  color(~Prediction > 0 & LowerCI > 0, ~Prediction, color="#539c4e") |>
  bold(~Prediction > 0 & LowerCI > 0, ~Prediction, bold=TRUE) |>
  color(~Prediction < 0 & UpperCI < 0, ~Prediction, color="#ad2323") |>
  bold(~Prediction < 0 & UpperCI < 0, ~Prediction, bold=TRUE) |>
  bold(part="header", bold=TRUE) |>
  set_table_properties(layout = "autofit") |>
  bg(bg="lightgrey", part = "header") |>
  fontsize(size = 9, part = "body") |>
  set_formatter(Prediction = function(x) {
    formatC(x, format = "e", digits = 1)
  }) |>
  set_formatter(LowerCI = function(x) {
    formatC(x, format = "e", digits = 1)
  }) |>
    set_formatter(UpperCI = function(x) {
    formatC(x, format = "e", digits = 1)
  })

table_pred

if(file.exists(here("table_nipagin_effect_Pe.docx")))
{
}else{
    save_as_docx(table_pred, path = here("table_nipagin_effect_Pe.docx"))
  }


table_pred2 <-flextable(fits_bay1 |>
  mutate(.iteration = .draw) |>
  group_by(treatment) |>
  compare_levels(exp.epred, by = time, comparison = control) |>
  mean_hdi(exp.epred) |> ## here with mean, but it also works with mean_hdi, median_qi, mean_qi
  mutate(.lower = round(.lower, digits = 2)) |>
  mutate(.upper = round(.upper, digits = 2)) |>
  mutate(exp.epred = round(exp.epred, digits = 2)) |>
  dplyr::select(treatment, time, exp.epred, .lower, .upper) |>
  rename("LowerCI" = ".lower", "Prediction" = "exp.epred", "UpperCI" = ".upper", "Treatment" = "treatment", "Time" = "time")) |>
  color(~Prediction > 0 & LowerCI > 0, ~Prediction, color="#539c4e") |>
  bold(~Prediction > 0 & LowerCI > 0, ~Prediction, bold=TRUE) |>
  color(~Prediction < 0 & UpperCI < 0, ~Prediction, color="#ad2323") |>
  bold(~Prediction < 0 & UpperCI < 0, ~Prediction, bold=TRUE) |>
  bold(part="header", bold=TRUE) |>
  set_table_properties(layout = "autofit") |>
  bg(bg="lightgrey", part = "header") |>
  fontsize(size = 9, part = "body") |>
  set_formatter(Prediction = function(x) {
    formatC(x, format = "e", digits = 1)
  }) |>
  set_formatter(LowerCI = function(x) {
    formatC(x, format = "e", digits = 1)
  }) |>
    set_formatter(UpperCI = function(x) {
    formatC(x, format = "e", digits = 1)
  })

table_pred2

if(file.exists(here("table_nipagin_effect_Pe2.docx")))
{
}else{
    save_as_docx(table_pred2, path = here("table_nipagin_effect_Pe2.docx"))
  }

```